---
title: "Deliverable 1: Project Plan"
author: "Practicum Spring 2024"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: style.css
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Workforce Housing in Watauga County, NC

## Project Scope

Watauga County, North Carolina, is currently experiencing a housing
shortage that is expected to intensify in the near future. The shortage
is exacerbated by vacation rentals and student housing, which has
significantly reduced the available residential housing options.
Currently, there is no streamlined system in place to help identify
suitable parcels for development. In response to this issue, the Watauga
Housing Council has contracted us to develop an application-based data
framework that is designed to identify land parcels based on their
likelihood to be developed for affordable and workforce housing. This
data framework will be designed to function as an interactive
application interface that will allow users to input criteria related to
land supply and combine it with development likelihood estimates to rank
each parcel in the county in terms of suitability. We will use the
administrative record associated with a septic system permit as the key
indicator for development. This permit is exclusively granted when a
property satisfies the requirements for both land supply and developer
demand at that specific site. Our approach will involve identifying the
site suitability characteristics (e.g., slope, soils) most closely
related to the probability of a parcel receiving a septic permit.
Subsequently, we will forecast the likelihood that a given parcel will
receive a new septic permit application in the next ten years based on
the analysis of the previous ten years of permits.

## Project Objectives

-   Identify the features that influence septic permit receival.

-   Create a model that can predict to a degree of certainty which
    parcels will receive a septic permit.

-   Create a data framework that identifies parcels based on their
    estimated likelihood of being developed.

-   Create a interactive user interface application where the user can
    search land parcels based on input criteria related to land supply.

*Goals can be further broken down into phases:*

**Exploratory Analysis**

-   Understand the current state of housing and affordability,
    development trends, and demographics in Watauga and surrounding
    countries.

-   Identify the characteristics of properties that are sold in the
    County.

-   Data wrangling -- Collect data that influences.

-   Collect septic system permit data and Identify the features that
    influence the approval process.

**Modeling**

-   Test and finalize model predictors.

-   Figure out temporal construction of the model.

-   Validate model.

**App Development**

-   Create app wireframe.

-   Develop app.

-   Test app function.

## Project Schedule

```{r gant, echo=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(Cairo)

tasks <- read.csv("tasks.csv")

tasks.long <- tasks %>%
  mutate(Start = mdy(Start),
         End = mdy(End)) %>%
  gather(date.type, task.date, -c(Project, Task)) %>%
  arrange(date.type, task.date) %>%
  mutate(Task = factor(Task, levels=rev(unique(Task)), ordered=TRUE))

# Custom theme for making a clean Gantt chart
theme_gantt <- function(base_size=11, base_family="Arial") {
  ret <- theme_bw(base_size, base_family) %+replace%
    theme(panel.background = element_rect(fill="#ffffff", colour=NA),
          axis.title.x=element_text(vjust=-0.2), axis.title.y=element_text(vjust=1.5),
          title=element_text(vjust=1.2, family="Arial"),
          panel.border = element_blank(), axis.line=element_blank(),
          panel.grid.minor=element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(size=0.5, colour="grey80"),
          axis.ticks=element_blank(),
          legend.position="bottom", 
          axis.title=element_text(size=rel(0.8), family="Arial"),
          strip.text=element_text(size=rel(1), family="Arial"),
          strip.background=element_rect(fill="#ffffff", colour=NA),
          panel.spacing.y=unit(1.5, "lines"),
          legend.key = element_blank())
  
  ret
}

custom_palette <- c("#8c0101",  "#cc621b", "#ccab18",  "#bcd15e", "#135418", "#a6e0c5","#4287f5", "#142996", "#8e13b0")


# Calculate where to put the dotted lines that show up every three entries
x.breaks <- seq(length(tasks$Task) + 0.5 - 3, 0, by=-3)

# Build plot with adjusted legend placement
timeline <- ggplot(tasks.long, aes(x=Task, y=task.date, colour=Project)) + 
  geom_line(size=3, width = 0.6) + 
  geom_vline(xintercept=x.breaks, colour="grey80", linetype="dotted") + 
  scale_color_manual(values = custom_palette) +
  guides(colour=guide_legend(title=NULL)) +
  labs(x=NULL, y=NULL) + coord_flip() +
  scale_y_date(date_breaks="1 week", date_labels = "%d %b") +
  theme_gantt() + 
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position = "right")  # Adjust the position of the legend

# Plot the timeline
timeline

```

## Risk Management

Potential risks we may encounter include:

-   Issues with code and modeling (optimization, validation, etc.)

-   Issues with app development and user interface

-   Possible issues with knitting

## Communication Plan

**Internal communication**: Whatsapp/Email for messaging. Plan to meet
on Thursdays outside of class for collaborative working hours.

**Client communication**: initial meeting + interview held over zoom,
other communication largely held over email.

## Stakeholder Management

Watauga Housing Council: (Kellie Reed Ashcraft, Laura Beach, Chris
Quattro (Appalachian State University)): Main project client, this
organization is tasked with responding to the economic, environmental,
and social aspects of the county to provide adequate housing options for
all residents. AppHealthCare (Andrew Blethen, Jason Pierce): besides
their widely-known health-related services, this organization is also
tasked with operating the permitting of septic and water infrastructure
in the County.

## Change Management

Because we developed an initial project plan, complete with deadlines,
deliverables, and tasks, it will be more clear and opaque on how to
address updates that may come up. Each week, we will evaluate our
progress against the Project Plan, readjusting deadlines and aspects of
the deliverables as appropriate. Maintaining flexibility while staying
on task will be crucial to a successful project.

## Team Roles + Responsibilities

**Project Managers** (*Jarred, Kathleen*) - Keeps team members on task
and accountable Leads team sprints Manages communications

**Markdown + Github Lead**\* - (*Jingyi, Kathleen*) - Keeps team
markdown updated Publishes md / rmd for professor review each week

**JS Engineering Lead** - (*Alice, Yinan*) - Leads app development

**Model Engineering Lead** - *Jingyi*

**Lead Researcher** - *Jarred*

**Exploratory Analysis Lead** - *Yinan*

**Lead Writer** - *Kathleen*

## Monitoring + Evaluation

We are in the process of developing a comprehensive project task list,
which will include a detailed breakdown of the project scope, outlining
all major objectives and their corresponding subtasks. We will conduct
weekly review meetings where we will discuss the progress of each task
in conjunction with the project schedule. This will ensure that we
remain on track with the projected deadlines and allows us to make any
necessary adjustments to keep the project aligned with our goals.

## Closure + Documentation

We will develop a project closeout checklist that will outline all
necessary steps for creating the final deliverables and documentation.
Once the checklist is finalized, we will assign specific tasks to each
team member and collaborate to ensure the completion of all necessary
activities. Additionally, we have allocated a sufficient amount of time
to facilitate the development and completion of all tasks involved in
the project closeout process.
