---
title: "Distance_Processing"
output: html_document
date: "2024-02-19"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE , results='hide')
```

# Data Loading

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message = FALSE, warning = FALSE)

library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(classInt)   # for KDE and ML risk class intervals
library(mapview)
# functions
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

```


```{r pressure, echo=FALSE}
landslide = st_read("https://raw.githubusercontent.com/JarredERandall/Musa_801_Watauga/main/Data/natural/landslide.geojson")

boundary = st_read('https://raw.githubusercontent.com/JarredERandall/Musa_801_Watauga/main/Data/natural/Boundary.geojson')

watersupply = st_read('https://raw.githubusercontent.com/JarredERandall/Musa_801_Watauga/main/Data/natural/water_supply.geojson')
```


```{r pressure, echo=FALSE}
line = st_read('https://raw.githubusercontent.com/JarredERandall/Musa_801_Watauga/main/Data/natural/line.geojson')

parcel = st_read('https://raw.githubusercontent.com/JarredERandall/Musa_801_Watauga/main/Data/natural/parcel.geojson')
```

```{r}
# Calculate the centroid of each parcel
parcel_centroids <- st_centroid(parcel)
```


```{r}
distance = parcel

# Add a column named 'centroid_geometry' to the parcel dataset
distance$centroid_geometry <- parcel_centroids
```


# Distance to landslide
```{r}
landslide_geometry <- st_geometry(landslide)

# Convert the geometry to a data frame
landslide_df <- as.data.frame(landslide_geometry)
```

```{r}
st_c <- st_coordinates
st_coid <- st_centroid

test1<-st_c(st_coid(parcel))
test2<- st_c(st_coid(landslide))
test3 <- st_c(st_coid(watersupply))

final <- parcel %>%
    mutate(
      landslide.nn =
        nn_function(test1, test2,k=2),
      watershed.nn = 
        nn_function(test1, test3, k = 1))
```


```{r}
ggplot() +
      geom_sf(data = final, aes(fill=landslide.nn), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Landslide Distance") +
      labs(title="Distance to the two Nearest Landslide Points") +
      mapTheme()
```

Landslides pose the primary hazard in Watauga County, and the safety of each parcel correlates with its distance from landslide-prone areas. Illustrated in the provided figure, the northeastern region of Watauga emerges as the safest zone. This safety is partly attributable to the gentle slope characteristic of this particular area.

```{r save}
plot <-ggplot() +
      geom_sf(data = final, aes(fill=landslide.nn), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Landslide Distance") +
      labs(title="Distance to the Two Nearest Landslide Points") +
      mapTheme()
ggsave("landslide.jpg", plot, dpi = 300)
```

# Distance to watershed

```{r}
ggplot() +
      geom_sf(data = final, aes(fill=watershed.nn), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Watersupply Distance") +
      labs(title="Distance to the Nearest Water supply") +
      mapTheme()
```
Water supply plays a crucial role in Watauga County's landscape. Given the potential environmental ramifications, it's imperative to maintain a safe distance between the septic systems and the water supply areas. Through comprehensive analysis, it's evident that the middle region of Watauga County is adjacent to these critical water supply zones. Consequently, when installing septic systems in this area, careful attention must be paid to mitigating environmental risks and preventing water pollution.


```{r}
plot <-ggplot() +
      geom_sf(data = final, aes(fill=watershed.nn), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Watersupply Distance") +
      labs(title="Distance to the Nearest Water supply") +
      mapTheme()
ggsave("watersupply.jpg", plot, dpi = 300)

```

# Line_Distance

```{r}
line <- line %>%
               rename(Dist_Road = `Road_Dist`)
```

```{r}
line2 <- st_transform(line, st_crs(final))
combined_df <- cbind(final, line2)
```


## Distance to nearest Road
```{r}
ggplot() +
      geom_sf(data = combined_df, aes(fill=Dist_Road), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Road Distance") +
      labs(title="Distance to the Nearest Road") +
      mapTheme()
```
Accessibility to transportation infrastructure significantly influences the development potential of a parcel. Typically, proximity to the road network directly correlates with higher accessibility. However, the graphic reveals certain parcels lacking adequate access to roads, indicating potential limitations in their development value. After analysis we can see that south-west and north-east Watauga is less connected.

```{r}
plot <-ggplot() +
      geom_sf(data = combined_df, aes(fill=Dist_Road), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Road Distance") +
      labs(title="Distance to the Nearest Road") +
      mapTheme()
ggsave("road.jpg", plot, dpi = 300)

```


## Distance to nearest Waterbody

```{r}
ggplot() +
      geom_sf(data = combined_df, aes(fill=Dist_Flowline), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Water Distance") +
      labs(title="Distance to the Nearest Waterbody") +
      mapTheme()
```

The presence of water bodies is integral to our evaluation framework. They significantly influence the water table, consequently affecting the design and placement of septic systems. Watauga County boasts a dense network of waterways. Despite this, we prioritize calculating the distance from the center of mass to the nearest river in our assessments. Overall, the variance in proximity to water bodies among parcels is minimal.

```{r}
plot <-ggplot() +
      geom_sf(data = combined_df, aes(fill=Dist_Flowline), colour=NA) +
  #geom_sf(data = parcel, fill = "transparent", colour = "red", size = 0.005) +  # Adjust the size here
      scale_fill_viridis(name="Water Distance") +
      labs(title="Distance to the Nearest Waterbody") +
      mapTheme()
ggsave("waterbody.jpg", plot, dpi = 300)

```